Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators

### 2025-05-08 research_proposal/research_proposal_draft.qmd
Pseudo algorithm

---
title: "Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators"
format: 
  pdf:
    include-in-header:
      text: |
        <!-- \usepackage{amsmath} -->
        <!-- \usepackage{mathspec} -->
        <!-- \usepackage{algorithm} -->
        <!-- \usepackage{algpseudocode} -->
        \usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
    mathspec: true
    number-sections: true
execute:
  cache: true
bibliography: D:/Github/Recon_Honours_Thesis/references.bib
# bibliographystyle: apa
csl: D:/Github/Recon_Honours_Thesis/apa.csl
---

@@

We reconstruct the estimator of $\boldsymbol{W}_h$ as $\hat{\boldsymbol{W}}_h = k_h \, g(\hat{\boldsymbol{W}}_1)$, where $k_h > 0$. The function $g(.)$ is an estimator of the unbiased sample covariance matrix of the in-sample one-step-ahead base forecast errors $\hat{\boldsymbol{W}}_1 = \frac{1}{T} \sum_{t=1}^{T} \hat{\boldsymbol{e}}_{t|t-1} \hat{\boldsymbol{e}}_{t|t-1}'$, and $T$ is the length of series' time dimension.

@@

$$
\hat{\boldsymbol{W}}^{N}_{1} = \lambda_{\delta} \hat{\boldsymbol{W}}_{1, \delta} + (1 - \lambda_{\delta}) \hat{\boldsymbol{W}}_1,
$$

@@

$$
\hat{\boldsymbol{R}}^{N}_{1} = \lambda_{\delta} \hat{\boldsymbol{R}}_{1,\delta} + (1 - \lambda_{\delta}) \hat{\boldsymbol{R}}_1,
$$

@@

$$
\hat{r}_{1,ij}^\delta = \text{sign}(\hat{r}_{1,ij}) \, 
\text{max}(|\hat{r}_{1,ij}| - \delta, \; 0),
$$

@@

$$
\hat{\lambda}(\delta) = \frac{
  \sum_{i \neq j} \widehat{Var}(\hat{r}_{1,ij}) \; \boldsymbol{1}(|\hat{r}_{1,ij}| \leq \delta)
} {
  \sum_{i \neq j} (\hat{r}_{1,ij} - \hat{r}_{1,ij}^\delta)^2
},
$$

@@

![NOVELIST cross-validation procedure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/novelist_cv_algo.png){#fig-novelist_cv height=70%}

@@

<!-- ```{=latex} -->
<!-- \begin{algorithm} -->
<!-- \caption{Cross-validation procedure to select optimal threshold $\delta^*$ for NOVELIST} -->
<!-- \begin{algorithmic}[1] -->
<!-- \State \textbf{Input:} Residual matrix $E$ of dimension $T \times p$, grid of threshold candidates $\{\delta_j\}$, window size $n$ -->
<!-- \For{each threshold candidate $\delta_j$} -->
<!--     \For{$t = n+1$ to $T$} -->
<!--         \State Define in-sample window: $E_{1:n} = \{e_{t-n}, ..., e_{t-1}\}$ -->
<!--         \State Compute sample covariance $\hat{\Sigma}_t = \frac{1}{n} E_{1:n}^\top E_{1:n}$ -->
<!--         \State Compute thresholded covariance: $\hat{\Sigma}^{\text{thr}}_t = \text{Threshold}(\hat{\Sigma}_t, \delta_j)$ -->
<!--         \State Compute NOVELIST: $\hat{\Sigma}^{\text{nov}}_t = (1 - \lambda) \hat{\Sigma}_t + \lambda \hat{\Sigma}^{\text{thr}}_t$ -->
<!--         \State Compute inverse: $W_t = (\hat{\Sigma}^{\text{nov}}_t)^{-1}$ -->
<!--         \State Compute MinT reconciliation matrix: $P_t = (S^\top W_t S)^{-1} S^\top W_t$ -->
<!--         \State Compute forecast error: $e_t = y_t - S P_t \hat{y}_t$ -->
<!--     \EndFor -->
<!--     \State Compute total CV loss $L_j = \sum_t \|e_t\|^2$ -->
<!-- \EndFor -->
<!-- \State \textbf{Output:} $\delta^* = \arg\min_{\delta_j} L_j$ -->
<!-- \end{algorithmic} -->
<!-- \end{algorithm} -->
<!-- ``` -->

>>>>

---
title: "Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators"
format: 
  pdf:
    include-in-header:
      text: |
        <!-- \usepackage{amsmath} -->
        <!-- \usepackage{mathspec} -->
        \usepackage{algorithm}
        \usepackage{algpseudocode}
        \usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
    mathspec: true
    number-sections: true
execute:
  cache: true
bibliography: D:/Github/Recon_Honours_Thesis/references.bib
# bibliographystyle: apa
csl: D:/Github/Recon_Honours_Thesis/apa.csl
---

@@

We reconstruct the estimator of $\boldsymbol{W}_h$ as $\hat{\boldsymbol{W}}_h = k_h \, g(\hat{\boldsymbol{W}}_1)$, where $k_h > 0$. The function $g(.)$ is an estimator of the unbiased sample covariance matrix of the in-sample one-step-ahead base forecast errors $\hat{\boldsymbol{W}}_1 = \frac{1}{T} \sum_{t=1}^{T} \hat{\boldsymbol{e}}_{t} \hat{\boldsymbol{e}}_{t}'$, and $T$ is the length of series' time dimension.

<!-- $\boldsymbol{e}}_{t|t-1}$ we keep things simplified without the conditional subscript -->

@@

$$
\hat{\boldsymbol{W}}^{N}_{1} = \lambda_{\delta} \hat{\boldsymbol{W}}_{1, \delta} + (1 - \lambda_{\delta}) \hat{\boldsymbol{W}}_1,
$$ {#eq-novelist-cov}

@@

$$
\hat{\boldsymbol{R}}^{N}_{1} = \lambda_{\delta} \hat{\boldsymbol{R}}_{1,\delta} + (1 - \lambda_{\delta}) \hat{\boldsymbol{R}}_1,
$$ {#eq-novelist-cor}

@@

$$
\hat{r}_{1,ij}^\delta = \text{sign}(\hat{r}_{1,ij}) \, 
\text{max}(|\hat{r}_{1,ij}| - \delta, \; 0),
$$ {#eq-soft-thr}

@@

$$
\hat{\lambda}(\delta) = \frac{
  \sum_{i \neq j} \widehat{Var}(\hat{r}_{1,ij}) \; \boldsymbol{1}(|\hat{r}_{1,ij}| \leq \delta)
} {
  \sum_{i \neq j} (\hat{r}_{1,ij} - \hat{r}_{1,ij}^\delta)^2
},
$$ {#eq-lambda-thr}

@@

```{=latex}
\begin{algorithm}
\caption{Cross-validation procedure}
\begin{algorithmic}[1]

\State \textbf{Input:} Observations and fitted values $\boldsymbol{y}_t, \hat{\boldsymbol{y}}_t \in \mathbb{R}^n$ for $t = 1,\dots,T$, set of threshold candidates $\Delta$, window size $v$.

\State $\hat{\boldsymbol{e}}_t = \boldsymbol{y}_t - \hat{\boldsymbol{y}}_t$ for $t = 1,\dots,T$

\For{$i = v:T-1$ do}
    \State $j = i - v +1$
    \State $\hat{\boldsymbol{W}}_j = \frac{1}{v} \sum_{t=j}^{i} \hat{\boldsymbol{e}}_{t} \hat{\boldsymbol{e}}_{t}'$
    \State $\hat{\boldsymbol{D}}_j = \text{diag}(\hat{\boldsymbol{W}}_j)$
    \State $\hat{\boldsymbol{R}}_j = \hat{\boldsymbol{D}}_j^{-1/2} \hat{\boldsymbol{W}}_j \hat{\boldsymbol{D}}_j^{-1/2}$
    \For{$\delta \in \Delta$}
        \State Compute thresholded correlation $\hat{\boldsymbol{R}}_{j,\delta}$ using Equation 5
        \State Compute $\hat{\lambda}_{j,\delta}$ using Equation 6
        \State Compute $\hat{\boldsymbol{R}}^{N}_{j,\delta}$ using Equation 4
        \State $\hat{\boldsymbol{W}}^{N}_{j,\delta} = \hat{\boldsymbol{D}}_j^{1/2} \hat{\boldsymbol{R}}^{N}_{j,\delta} \hat{\boldsymbol{D}}_j^{1/2}$
        \State $\boldsymbol{G} = (\boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{j,\delta} \boldsymbol{S})^{-1} \boldsymbol{S}' \hat{\boldsymbol{W}}^{N^{-1}}_{j,\delta}$
        \State Reconciled forecasts $\tilde{\boldsymbol{y}}_{i+1 | \delta} = \boldsymbol{S} \boldsymbol{G} \hat{\boldsymbol{y}}_{i+1}$
        \State $\tilde{\boldsymbol{e}}_{i+1 | \delta} = \boldsymbol{y}_{i+1} - \tilde{\boldsymbol{y}}_{i+1 | \delta}$
    \EndFor
\EndFor

\State $\text{MSE}_{\delta} = \frac{1}{T-v} \sum_{i=v}^{T-1} (\tilde{\boldsymbol{e}}_{i+1 | \delta})^2$ for each $\delta \in \Delta$
\State $\hat{\delta}^* = \arg\min_{\delta \in \Delta} \text{MSE}_{\delta}$
\State Compute $\hat{\lambda}^*$ on all training data using $\hat{\delta}^*$
\State Compute $\hat{\boldsymbol{R}}_1^*$ using $\hat{\delta}^*$ and $\hat{\lambda}^*$ on all training data, using Equation 3

\State \textbf{Output:} Estimate of optimal $\hat{\delta}^*$

\end{algorithmic}
\end{algorithm}
```

<!-- ![NOVELIST cross-validation procedure](D:/Github/Recon_Honours_Thesis/research_proposal/figs/novelist_cv_algo.png){#fig-novelist_cv height=70%} -->

@@
